### 2020-02-26 thesis/paper.tex
edit

To visualize how similar different clusterings are one needs to define a measure representing the distance between two results. Using this measure, finding groups of similar results is interesting as those may indicate robust solutions. Still, it is difficult to define how close clustering results need to be for them to be considered similar. For this reason, I use OPTICS for meat-clustering the results as it produces a reachability plot in which a user can visually identify how this similarity could be defined. This can be done, as the plot shows valleys for dense regions of objects (in this case clusterings) and these valleys are quite simple to see.

\subsection{Gathering more information}
Another research question was: How can we gain knowledge from more computed clusterings, instead of just picking one? I found that the goal of most visual frameworks is to find the best clustering solution out of a bunch of computed ones. Using the information of more than one clustering can result in a better solution though, so finding a way of including it was relevant. Here, this was achieved by using consensus functions which are able to combine multiple clustering solutions into one. This can not only be done by combining all computed solutions but in combination with finding similar groups, these groups can be combined into one result each.

Also, the question of how one can obtain candidate solutions was of importance. In many data-sets the number of clusters or the structure of them may not be unambiguously defined. It may be the case, that a group of points has a gap in the data due to random factors or that this group actually consists of two groups. It would now be interesting for a user to find which solutions are possible and to reason about why one solution may be better than another one. To solve this, the combination of similar clusterings is useful as each combined result shows the consensus of that meta-cluster. If the meta-cluster was large and defined by how closely related the results were, its size can be used as an indication for the combined solution being satisfactory. This can be assumed, as the user can reason that for the different obtained solutions from different meta-clusters, the ones from large ones are more robust than the ones from small ones.

\subsection{Easy Usage of complex Methods/Workflows}
The aspect of how easily a user can use those often complex clustering and consensus clustering methods was also relevant. There are currently no simple tools that allow using clustering methods, compare them and compute consensus clusterings from them, all in one place. Therefore, an important consideration was to make this process as easy as possible, while not loosing out on the power the methods bring with them. The strategy here was to clearly separate the pre-processing, defining of the clustering algorithms and the exploration and merging of the results. To also enable the use of future or currently missing methods, the extendability of the tool was of importance, which is achieve through interfaces with which such methods can easily be added.

>>>>

To visualize how similar different clusterings are one needs to define a measure representing the distance between two results. Using this measure, finding groups of similar results is interesting as those may indicate robust solutions. Still, it is difficult to define how close clustering results need to be for them to be considered similar. For this reason, I use OPTICS for meta-clustering the results as it produces a reachability plot in which a user can visually identify how this similarity could be defined. This can be done, as the plot shows valleys for dense regions of objects (in this case clusterings) and these valleys are quite simple to see. With the ordering produced by OPTICS, a heat map of the distance matrix of the clusterings can also be shown, there the actual distances can be more finely explored. Also, a projection to points can be shown for this matrix by using MDS, this can give a fast approximate overview of which solutions are close to each other.

\subsection{Gathering more Information}
Another research question was: How can we gain knowledge from more computed clusterings, instead of just picking one? I found that the goal of most visual frameworks is to find the best clustering solution out of a bunch of computed ones. Using the information of more than one clustering can result in a better solution though, so finding a way of including it was relevant. Here, this was achieved by using consensus functions which are able to combine multiple clustering solutions into one. This can not only be done by combining all computed solutions but also in combination with finding similar groups. These groups can be combined into one result each, capturing the essence of all included results.

Also, the question of how one can obtain multiple candidate solutions was of importance. In many data-sets the number of clusters or the structure of them may not be unambiguously defined. It may be the case, that a group of points has a gap in the data due to random factors or that this group actually consists of two groups. It would now be interesting for a user to find which solutions are possible and to reason about why one solution may be better than another one. To solve this, the combination of similar clusterings is useful as each combined result shows the consensus of that meta-cluster. If the meta-cluster was large and defined by how closely related the results were, its size can be used as an indication for the combined solution being satisfactory. This can be assumed as the user can reason that for the different obtained solutions from different meta-clusters, the ones from large ones are more robust than the ones from small meta-clusters.

\subsection{Easy Usage of complex Methods/Workflows}
The aspect of how easily a user can use those often complex clustering and consensus clustering methods was also relevant. There are currently no simple tools that allow to use clustering methods, compare them and compute consensus clusterings from them, all in one place. Therefore, an important consideration was to make this process as easy as possible, while not loosing out on the power the methods bring with them. The strategy here was to clearly separate the pre-processing, the definition of the clustering algorithms and the exploration and merging of the results. To also enable the use of future or currently missing methods, the extendability of the tool was of importance, this is achieved through interfaces with which such methods can easily be added.
